---
title: "Data Availability"
description: |
  On this page, you can find information on obtaining sequence data, data products, and processing scripts. Code is  embedded in indivudual workflows on the website or can be accessed through the project GitHub repo.
bibliography: assets/cite.bib  
---

```{r, xaringanExtra-clipboard, echo=FALSE, eval=TRUE}
htmltools::tagList(
  xaringanExtra::use_clipboard(
    button_text = "<i class=\"fa fa-clone fa-2x\" style=\"color: #301e64\"></i>",
    success_text = "<i class=\"fa fa-check fa-2x\" style=\"color: #90BE6D\"></i>",
    error_text = "<i class=\"fa fa-times fa-2x\" style=\"color: #F94144\"></i>"
  ),
  rmarkdown::html_dependency_font_awesome()
)
```

## Archived Sequence Data

Raw **16S rRNA** data for each sample (before removing primers) is available on figshare at [doi:10.25573/data.14686665](https://doi.org/10.25573/data.14686665). Raw **ITS** data for each sample (before removing primers) is available on figshare at [doi:10.25573/data.14686755](https://doi.org/10.25573/data.14686755). 

All trimmed sequence data (primers removed) is deposited at the European Nucleotide Archive (ENA) under the study accession number [PRJEB45074 (ERP129199)](https://www.ebi.ac.uk/ena/browser/view/PRJEB45074). The trimmed **16S rRNA** data (primers removed) are deposited under sample accession numbers [ERS6485270-ERS6485284](https://www.ebi.ac.uk/ena/browser/view/ERS6485270-ERS6485284). The trimmed **ITS** data (primers removed) are deposited under sample accession numbers [ERS6485285-ERS6485299](https://www.ebi.ac.uk/ena/browser/view/ERS6485285-ERS6485299).

## Pipeline Data

Data for each individual pipeline are available through the Smithsonian figshare under a single collection at [doi:10.25573/data.c.5667571.v2](https://doi.org/10.25573/data.c.5667571.v2). In addition, data from each pipeline are available for download from figshare using the links at the bottom of each workflow page (where applicable).

<details markdown="1">
<summary>**Show** descriptions and DOI's for each workflow</summary>

</br>

Scott, Jarrod (2022): (high-temp/16s) No 0. Raw 16S rRNA fastq data files for the high-temp SWELTR study. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.14686665.v1

Scott, Jarrod (2022): (high-temp/ITS) No 0. Raw ITS fastq data files for the high-temp SWELTR study. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.14686755.v1

Scott, Jarrod (2022): (high-temp) No 1. DADA2 Workflow (16S rRNA/ITS) Output. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.14687184.v1

Scott, Jarrod (2022): (high-temp) No 2. Data Preparation (16S rRNA/ITS) Output. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.14690739.v1

Scott, Jarrod (2022): (high-temp) No 3. Filtering: (16S rRNA/ITS) Output. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.14701440.v1

Scott, Jarrod (2022): (high-temp) No 4. Taxonomic: (16S rRNA/ITS) Output. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.16826737.v1

Scott, Jarrod (2022): (high-temp) No 5. Aplha diversity (16S rRNA/ITS) Output. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.16826779.v1

Scott, Jarrod (2022): (high-temp) No 6. Beta diversity (16S rRNA/ITS) Output. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.16828063.v1

Scott, Jarrod (2022): (high-temp) No 7. Differentially Abundant ASVs & Taxa (16S rRNA/ITS) Output. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.16828249.v1

Scott, Jarrod (2022): (high-temp) No 8. Metadata Analysis (16S rRNA/ITS) Output. Smithsonian Tropical Research Institute. Dataset. https://doi.org/10.25573/data.16828294.v1

</details>

## Submitting Sequence Data

We submitted out data to the [European Nucleotide Archive (ENA)](https://www.ebi.ac.uk/ena). The ENA does not like RAW data and prefers to have primers removed. So we submitted the trimmed Fastq files to the ENA. You can find these data under the study accession number **PRJEB45074 (ERP129199)**. The RAW files on our figshare site (see above).

To submit to the ENA you need two data tables (plus your sequence data). One file describes the samples and the other file describes the sequencing data.

You can download our submission data tables here:

* [Sample submission form](include/data-availability/ENA_Sample_SUBMISSION_SWELTR-high-temp.tsv)
* [FASTq submission form](include/data-availability/ENA_PAIRED_FASTQ_SUBMISSION_SWELTR-high-temp.tsv)

Note, these forms are study specific, so please use these as guides only.

### Instructions for Submitting to the ENA

Even though I have done this dozens of times, the process of submitting sequence data to read archives still baffles me. 

> Note: I submit data with primers, barcodes, etc **removed**

#### Register Project & Upload Sample Data

1) go to https://www.ebi.ac.uk/ena/submit and select **Submit to ENA Interactively**.
2) Login or Register.
3) Go to **New Submission** tab and, if this is a new project, select **Register study (project)**.
4) Hit Next
5) Enter details and hit **Submit**.
6) Next, [Select Checklist](https://www.ebi.ac.uk/ena/submit/checklists). This will be specific to the type of samples you have and basically will create a template so you can add your sample metadata. For this study I chose **GSC MIxS soil**, checklist accession number [ERC000022](https://www.ebi.ac.uk/ena/data/view/ERC000022)
7) Next
8) Now go through and select/deselect fields as needed. Note, some fields are mandatory.
9) Once finished, hit **Next** to fill in any details that apply to *All* samples and the download the template. Alternatively, you can download the template and fill in the data by hand.
10) Upload the sample sheet. 
11) Once everything looks good and uploaded, click the **New Submission** tab.

#### Upload Sequence Data

12) Hit **Skip** and then select **Two Fastq files (Paired)**, Download the template and fill in the details.
13) Next, make sure all the trimmed fastq files are gzipped **.gz**  (these are what you submit). 
14) Navigate to the directory with  the trimmed, compressed fastq files and run:

```
md5sum *.gz
```

16) Add the checksums and file names to the fastq submission form. You can read more about [Preparing A File For Upload here](https://ena-docs.readthedocs.io/en/latest/submit/fileprep/preparation.html).

**STOP**

Before uploading the sheet, you first must upload the fastq files. Documentation for Uploading Files To ENA can be found [here](https://ena-docs.readthedocs.io/en/latest/submit/fileprep/upload.html). 

17) In the same directory, run:

```
lftp webin2.ebi.ac.uk -u Webin-XXXXX
```

where `Webin-XXXXX` is your user name. Enter your password

at the prompt run:

```
mput *.gz
# when finished
bye
```

The files should begin uploading. Depending on internet speed and/or file sizes/numbers, this could take a while. 

18) Once the upload is finished, upload and submit the fastq submission form. If the `sample alias` field dows not autofill you may need to upload the sample form again. 


## Source Code {.appendix}

The source code for this page can be accessed on GitHub by [clicking this link](https://github.com/sweltr/high-temp/blob/master/data-availability.Rmd).

## Last updated on {.appendix}

```{r,echo=FALSE, eval=TRUE}
Sys.time()
```
